# Week 5 Exercises"

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
```


Read the file timeuse_tidy.rds with `readRDS()`. The file contains the dataset that we tidied in the exercise session: records of daily time use from participants over multiple days. Note that since the data has been stored as rds (R-specific format), column types and factor levels are as we left them, and don't need to be re-corrected.

```{r}
tidytime <- readRDS("timeuse_tidy.rds")
glimpse(tidytime)
```


## Exercise 1

### 1.1

Create a new variable that contains combined activity classes: "Work or school" (Work, Business, Education), "Free time" (Shopping, Social / Leisure, Home, Vacation), and "Other".  

```{r}
act_key <- c("Work" = "Work or school", "Business" = "Work or school", "Education" = "Work or school", "Shopping" = "Free time", "Social / Leisure" = "Free time", "Home" = "Free time", "Vacation" = "Free time")

tidytime <- tidytime %>% mutate(activity_class = recode(activity_class, !!!act_key, .default = "Other") %>% factor)
```

Calculate the mean time spent on each of the combined activity classes, grouped by **weekday, participant ID, and occ_full_time**.

```{r}
mean_times <- tidytime %>% group_by(weekday, indivID, occ_full_time, activity_class) %>% summarise(mean(time_spent))
head(mean_times)
```

### 1.2 

Visualise the means you calculated.

```{r}
mean_times %>% ggplot(aes(x = activity_class, y = `mean(time_spent)`)) + 
  geom_violin(fill="#009bff") + # HY blue
  theme_bw() +
  scale_y_continuous(name = "Mean time spent") +
  scale_x_discrete(name = "Activity class")
```


## Exercise 2

### 2.1

What is computed in the code chunk below - what do the numbers tell you?  

Can you think of another way to calculate the same thing?  

```{r}

date_counts <- tidytime %>%
  distinct(indivID, date) %>% # Select distinct individuals and dates
  arrange(date) %>%           # Arrange the data by date
  count(date)                 # Count individuals per each date

date_counts                   # Show counts
```

### 2.2

Plot the numbers from above (use points, lines, or whatever you think is suitable).  

```{r}
date_counts %>% ggplot(aes(x = date, y = n)) +
  geom_point() +
  theme_minimal() +
  scale_y_continuous(name = "number of participants")
```


## Exercise 3

### 3.1

Count the **total number of participants** in the data.  

```{r}
tidytime %>% distinct(indivID) %>% count()
```

### 3.2

For each participant, count the **number of separate days** that they recorded their time use on.  

```{r}
tidytime %>% group_by(indivID) %>% distinct(date) %>% count(indivID)
```


## Exercise 4

Explain step by step what happens in the code chunk below and what the final figure represents.

_Explanations as comments. The final figure represents when the participants recorded their activities (points) with the overall all timespan (lines). The individuals are order by start date top to bottom and thus the figure shows this general trend left-to-right and top-to-bottom. Colors represent start month._

```{r fig.height=20, fig.width=10}
tidytime %>%
  group_by(indivID) %>% # Group by individuals 
  mutate(start_date = min(date)) %>% # Compute and add start date as minimum date per each individual (as it is grouped)
  ungroup %>% # ungroup the data
  mutate(indivID = factor(indivID), # factorize indivID variable
         indivID = fct_reorder(indivID, start_date) %>% fct_rev()) %>% # reorder indivID by start date
  ggplot(aes(date, indivID, colour = month(start_date, label = T))) + # init plot with date in x-axis and individuals on y-axis, color data by starting month that are displayed as a character strings
  geom_line() + # add line plot
  geom_point(size=.5, alpha=.1) + # add points and set point size and transparency (alpha channel for points' colors)
  theme_bw() + # set the bw-theme (dark-on-light)
  scale_y_discrete(breaks = "none") + # remove ticks form y-axis
  labs(x = "Date", y = "", colour = "Starting month") # set labels for the axes and 
```


## Extra - categorical time and value

I got intrested about Ben's example of restaurant menus as categorical time and value. And of course how to visualize those. Luckily, one restaurant/chef in Lohja, [Mikko Utter](http://www.mikkoutter.fi/kaudet.html), has menus for each season available at his webpage. So I crabbed those and coded the menu items very rough to food groups for simple  visualizations. Unfortunately, the seasons are in Finnish and I do not know good translations for all so I did not translate those.

```{r message=FALSE, warning=FALSE}
library(cowplot) #for side-by-side plots
```

Read the data `utter.csv´ and mutate to categories. Ids from row numbers that are then used to reorder the seasons so that they remain in temporal order.
```{r message=FALSE, warning=FALSE}
utter <- read_csv2("utter.csv", col_types = cols())

utter <- utter %>% mutate(
  season = as.factor(season),
  group = as.factor(group),
  id = as.numeric(row.names(utter))
)

utter <- utter %>% mutate(
  season = fct_reorder(season, id)
)

head(utter)
```
Simple bar plot and pie plot side-by-side. The pie plot is, according to the _Grammar of Graphics_, just a bar plot with polar coordinates. Thus making pie plot from bar plot with the ggplot is straightforward.
```{r, fig.width=10}
#Custom palette for nicer colors. Original from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/ but this is reordered to correspond bit nicer with the food groups
myPalette <- c("#56B4E9", "#E69F00", "#D55E00", "#F0E442", "#999999", "#CC79A7", "#0072B2", "#009E73")

p1 <- utter %>% ggplot(aes(x = season, fill = group)) + # Keep the order same as in the df
  geom_bar(width = 1) +
  scale_fill_manual(values=myPalette) +
  scale_y_discrete(breaks = NULL, expand = c(0, 0)) +
  labs(x="Seasons", y="") +
  theme_minimal() +
  theme(panel.grid = element_line(color = "white"))

# Wrap bars into a pie
start_angle = -0.5*(360/11)*(pi/180) # fixing start angle so that the first season "napa" is at the top
p2 <- p1 +coord_polar(start = start_angle)

plot_grid(p1, p2)
```
Usually pie plots are not recomended, but I like to visualize a year as circle so to me atleast the pie plot is more  intuitive and pleasing. Also, this video [Nightingale Diagrams - Numberphile](https://www.youtube.com/watch?v=VTdVPNvwULM) discusses and shows nice circular visualizations

Here is a horizontal bar plot of number of each food groups colored by the seasons. This uses the default color map of ggplot2 is quite horrendous, but then again I don't think this is a good visualization of this data.
```{r}
utter %>% ggplot(aes(y = group, fill = season)) +
  geom_bar(stat = "count", width = 1) +
  theme_classic()
```

Finally, let's make a pie plots of the grousp for each season. Again, I think here the pie plots are visually pleasing.
```{r, fig.width=8}
utter %>% group_by(season) %>% count(group) %>%  ggplot(aes(x = "", y = n, fill = group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_y_continuous(breaks = NULL) +
  facet_wrap(~season) +
  scale_fill_manual(values=myPalette) +
  labs(x="", y="") +
  theme_void()
```