# Week 3

```{r warning = FALSE, message = FALSE}
library(ggplot2)
library(tidyverse)
library(forcats)
```


## Exercise 1

### 1.1

Read the *region_scores.csv* data into R. It should have 240 observations of 6 variables.  

```{r message=FALSE}
df <- read_csv("region_scores.csv")
#Convert education to factorial
df <- df %>% mutate(education = as.factor(education))
df %>% summary #check summary of the data
```


### 1.2

Cut the continuous variable **age** into a categorical variable **age_group** that has groups of width 10. Use ggplot2's cutting function.  

Count observations by age group.

```{r}
df <- df %>% mutate(age_group = cut_width(age, 10))
df %>% count(age_group)
```


## Exercise 2

The chunk below is supposed to produce a plot but it has some errors.  

The figure should be a scatter plot of **age** on the x-axis and **score** on the y-axis, with points coloured by **education level** (3 levels). It should also have three linear regression lines, one for each of the education levels.  

Fix the code to produce the right figure.  

What happens if you use `geom_jitter()` instead of `geom_point()`? 

`Jitter` seems to give some random variation or noise. This is confirmed by ggplot documentation.

```{r warning=FALSE}
df %>% 
  ggplot(aes(x = age, y = score, color = education)) + geom_point() + 
  geom_smooth(method = "lm") +
  theme_light()

```


## Exercise 3

### 3.1

Calculate the mean, standard deviation (sd), and number of observations (n) of **score by education and age group**. Also calculate the standard error of the mean (by using sd and n). Save these into a new data frame (or tibble) named edu_age_stats.  

```{r}
edu_age_stats <- df %>% group_by(education, age_group) %>% summarise(mean_score=mean(score), sd_score=sd(score), n=n())
# Compute standard error of mean as sd/sqrt(n)
edu_age_stats$se_score = edu_age_stats$sd_score/sqrt(edu_age_stats$n)
```

### 3.2

Using edu_age_stats, plot a bar plot that has **age group** on the x-axis, **mean score** on the y-axis, and **education levels** in subplots (facets).  

Use `geom_errorbar()` to add error bars that represent standard errors of the mean.  

```{r}
edu_age_stats %>% ggplot(aes(x = age_group, y = mean_score)) +
  facet_wrap("education", labeller = "label_both") +
  geom_col(fill = "#fcd116") + # Color of our faculty for fun
  geom_errorbar(aes(ymin=mean_score-se_score, ymax=mean_score+se_score), colour = "#002b7f") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(name = "Mean score", expand = c(0, 0), limits = c(0, 20)) +
  scale_x_discrete(name = "Age Group")
```


## Exercise 4

### 4.1

Create a figure that has boxplots of **score** (x-axis) by **region** (y-axis).  

```{r}
df %>% ggplot(aes(x = score, y = region)) + geom_boxplot() + theme_classic()
```

### 4.2

Group the data by region and add mean score by region to a new column **mean_score**. Do this with **`mutate()`** (not `summarise()`). 

Reorder the levels of region based on mean_score.  
```{r}
df <- df %>% group_by(region) %>% mutate(mean_score=mean(score)) %>% ungroup()
df$region <- fct_reorder(df$region, df$mean_score, min)
```


## Exercise 5

Using the data you modified in exercise 4.2, plot **mean scores** (x-axis) by **region** (y-axis) as points. The regions should be ordered by mean score.  

Use `stat_summary()` to add error bars that represent standard errors of the mean.  
```{r}
df %>% ggplot(aes(mean_score, region)) +
  geom_point() + # All mean scores of each region at the same coordinates, geom_jitter would reveal this
  stat_summary(aes(score), fun.data = mean_se, geom="errorbar") + # Errorbars from scores
  theme_minimal() +
  scale_x_continuous(name = "Mean Score", breaks=c(10:20)) +
  scale_y_discrete(name = "Region")
```