# Week 2

## Exercise 1

### 1.1
Create a new code chunk where you load the tidyverse package. In the chunk settings, suppress any output messages.  
```{r setup, message=FALSE}
library(tidyverse)
```

### 1.2
The tibble *df* has 60 observations (rows) of variables (columns) group, gender, score1 and score2 (continuous scores from two tests). Each row represents one participant.  
```{r include = F}
# this chunk generates the data, and you can ignore it
set.seed(1)
df  <- tibble(group = sample(c(rep(1:2, each=30))),
              gender = sample(c(rep(c('M', 'F'), 28), rep(999,4))),
              score1 = rnorm(60, 18, 2),
              score2 = as.character(runif(60, 0.7, 1.1)*score1))
  
```

```{r}
df
```


There is something to fix in three of the variables. Explore the data and describe what needs to be corrected.  

**Hint**: You can use e.g. str(), distinct(), and summary() to explore the data.  
```{r}
str(df)
summary(df)
distinct(df, gender)
```
*group* could be categorical/factors, *gender* includes values of 999 and *score2* is character-type instead of numeric.

## Exercise 2

### 2.1

Make the corrections you described above.
```{r message=FALSE, warning=FALSE}
# Factorize group
df<- df %>% mutate(group = as.factor(group))

# Replace 999 with NA (not available)
df <- df %>% mutate(gender = na_if(gender, 999))

# Change score to numeric
df <- df %>% mutate(score2 = as.numeric(score2))

# Check structure again and distinct groups 
df %>% str()
df %>% distinct(gender)
```

### 2.2

Count observations by group and gender. Arrange by the number of observations (ascending).
```{r}
df %>% count(group, gender, name ="observations") %>% arrange(observations) # add column name for observations
```


## Exercise 3

### 3.1

Create a new variable, **score_diff**, that contains the difference between score1 and score2.

```{r}
# New variable score_diff into the dataframe with absolute value between score1 and score2
df$score_diff <- abs(df$score1-df$score2)
```

### 3.2

Compute the means of score1, score2, and score_diff.

**Hint**: Like mutate(), summarise() can take multiple variables in one go.

```{r}
df %>% summarise(score1_mean = mean(score1), score2_mean = mean(score2), score_diff_mean = mean(score_diff))
```

### 3.3

Compute the means of score1, score2, and score_diff **by gender**.

```{r}
df %>% group_by(gender) %>% summarise(mean(score1), mean(score2), mean(score_diff))
```


## Exercise 4

### 4.1

Using ggplot2, create a scatter plot with score1 on the x-axis and score2 on the y-axis. 
```{r}
df %>% ggplot(aes(x=score1, y=score2)) + geom_point()
```


### 4.2

Continuing with the previous plot, colour the points based on gender. 
Set the output figure width to 10 and height to 6.

```{r fig.width=10, fig.height=6}
df %>% ggplot(aes(x=score1, y=score2, color=gender)) + geom_point()
```

## Exercise 5

### 5.1

Add the author (your name) and date into the metadata section. Create a table of contents.  

### 5.2

Knit your document to HTML by changing `html_notebook` to `html_document` in the metadata, and pressing Knit.   

Did these for the returned ex-file!